---
title: "p8105_hw3_cev2129"
output: github_document
date: "2024-10-16"
---

```{r}

```

#Question 2
Loading the Nhanes_covar.csv 
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
covar_df = read_csv(file = "nhanes_covar.csv", skip = 4, col_names = TRUE, na = c(".", "NA", ""))
covar_df = drop_na(covar_df)
```
Loading Nhanes_accel.csv 
```{r}
accel_df = read_csv(file = "nhanes_accel.csv")
accel_df = drop_na(accel_df)
```
Joining the datasets
```{r}
nhanes_df = left_join(covar_df, accel_df, by = "SEQN", "age" > 21)
nhanes_df = nhanes_df |>
  mutate_all(as.character) 

```

Create a reader friendly table for number of men and women in each education category, and 
```{r}
nhanes_df |> 
  group_by(education, sex) |>      
  count() |>                      
  arrange(education)
#This table shows how many people from each sex has completed each level of education. Education level is shown twice so that it can be separated by sex. 
```


Create a visualization of the age distributions for men and women in each education category.
```{r}
nhanes_df = nhanes_df |>
  mutate(sex= as.factor(sex), education = as.factor(education), age = as.numeric(age))

ggplot(nhanes_df, aes(x= sex, y = age)) +   
  geom_boxplot() +           
  facet_wrap(~education)
#This box plot is meant to show what the reader friendly table depicts, howver you can more clearly view the distribtion of each sex in each education level. 
```


Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.
```{r}
nhanes_df <- nhanes_df |>
  mutate(total_activity = rowSums(across(min1:min1440, ~ as.numeric(.)), na.rm = TRUE))

ggplot(nhanes_df, aes(x = age, y = total_activity, color = sex)) + 
  geom_point() + geom_smooth() +
  facet_wrap(~ education)
#These plots show the differences in physical activity throughtout the day between those of different sex and ages, separated by education. Each level of education has a downward trend, however, those in the highest level of education seem to have the least steep slope. 
```


Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.
```{r}

```


#Question 3
```{r}
jan2020 = read_csv("Jan2020.csv")
jan2024 = read_csv("Jan2024.csv")
july2020 = read_csv("July2020.csv")
july2024 = read_csv("July2024.csv")
```